---
title: "Data selection"
author: "Sharaj"
format: html
editor: visual
---

Code to cleanup and select articles the raw dataframe for downstream annotation procedures and analysis

## Load libraries and dataframe

```{r}
library(stringr)

news <- read.csv("Dataset/Raw/newspaper_raw.csv")
```

## Cleanup

Here, articles were chosen for further analysis through the following steps:

1.  Articles that could not be scraped from newspaper4k were excluded.
2.  Articles that had fewer than 250 words were excluded.
3.  Article texts that didn't contain the term "monkeypox" or "mpox" were excluded.
4.  Articles that contained terms like "associated press", "wired" or "reuters" were removed to maximize diversity in the corpus and remove syndicated texts.
5.  Articles that contain more characters than the excel cell limit (32767) are excluded.

```{r}
#remove articles that we could not scrape
full <- which(news$text != "")
news_df <- news[full,]

#word count
word_count <- unlist(lapply(1:length(news_df$text), function(i) {
  str_count(news_df$text[i], '\\w+')
}))

#what is the head of the distribution
quantile(word_count, 0.1) #240

#mention of mpox or monkeypox
mpox_count <- unlist(lapply(1:length(news_df$text), function(i) {
  str_count(tolower(news_df$text[i]), 'monkeypox') + 
    str_count(tolower(news_df$text[i]), 'mpox')
}))

#mention of the associated press, reuters or wired
press_count <- unlist(lapply(1:length(news_df$text), function(i) {
  str_count(tolower(news_df$text[i]), 'associated press') + 
    str_count(tolower(news_df$text[i]), 'reuters') +
    str_count(tolower(news_df$text[i]), 'wired')
}))

#number of characters
char_count <- unlist(lapply(1:length(news_df$text), function(i) {
  nchar(tolower(news_df$text[i]))
}))

#exclusion criteria: 
keep <-  which((mpox_count>0)&(word_count>250)&(press_count==0)&(char_count<32767))

#choose certain columns
df1 <- news_df[keep,c("stories_id","media_name",
                      "title","url","text","author","keywords")]

df1 <- df1[order(df1$stories_id),]
rownames(df1) <- c(1:nrow(df1))

write.csv(df1, "Dataset/Pre-processing/mpox_news_data.csv", row.names = F)


```
