---
title: "Inter rater reliability between manual annotators"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Load libraries

```{r}
library(irr)
library(Metrics)
```

## Helper functions

### Calculating metrics

Create a function that outputs accuracy, cohen's kappa, f1 score, precision and recall between pairs of annotators:

1.  Standardized annotation dataframes for each annotator (dlist)
2.  The "theme" label (frame in this case) to compare annotations across
3.  Rater names A1 and A2
4.  Column names where the labels for A1 and A2 lie respectively
5.  Indicator on what to do with NA values (dropNA = T by default)

```{r}
calculate_scores <- function(dlist, A1, A2, column_A1, column_A2, column_match = "stories_id", dropNA = T)
{
  dfA1 <- dlist[[A1]]
  dfA2 <- dlist[[A2]]
  
  kdf1 <- merge(dfA1, dfA2, by.x = column_match, by.y = column_match, all = F)
  
  #if column names for A1 and A2 are same
  if(column_A1 == column_A2)
  {
    kdf <- kdf1[,c(paste0(column_A1, ".x"), paste0(column_A2, ".y"))]
  }
  
  else
  {
    kdf <- kdf1[,c(column_A1, column_A2)]
  }
  
  colnames(kdf) <- c("A1", "A2")
  
  if(dropNA)
  {
    kdf <- kdf[complete.cases(kdf),]
  }

  klist <- kappa2(kdf) #kappa calculation
  k <- round(klist$value, 2) #Cohen's Kappa
  
  acc <- round(accuracy(kdf[,1],kdf[,2]),2) #accuracy
  prec <- round(precision(kdf[,1], kdf[,2]), 2) #precision
  rec <- round(recall(kdf[,1], kdf[,2]), 2) #recall
  F1 <- round(2*prec*rec/(prec + rec), 2) #F1 score cos the f1 func is broken

  metric_vec <- c(acc, k, F1, prec, rec)
  print(paste0("Accuracy is ", acc, "; Cohen's kappa is ", k, "; F1 score is ",F1,
               "; Precision is ", prec, "; Recall is ", rec))
  
  return(metric_vec)
}
```

# Calculate scores

## Human - round 1

```{r}
df1 <- read.csv("Relevance/Human/A1_r1.csv")
df2 <- read.csv("Relevance/Human/A2_r1.csv")

dlist <- list("A1" = df1,
              "A2" = df2)

calculate_scores(dlist, A1 = "A1", A2 = "A2", 
                 column_A1 = "relevance", column_A2 = "relevance", column_match = "stories_id",
                 )

```

## Human - round 2

```{r}
df1 <- read.csv("Relevance/Human/A1_r2.csv")
df2 <- read.csv("Relevance/Human/A2_r2.csv")

dlist <- list("A1" = df1,
              "A2" = df2)

calculate_scores(dlist, A1 = "A1", A2 = "A2", 
                 column_A1 = "relevance", column_A2 = "relevance", column_match = "stories_id",
                 )

```
