---
title: "Formatting and saving labelled articles"
author: "Sharaj Kunjar"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

Code to compile all the labelled data

# Load libraries

```{r}
library(irr)
library(Metrics)
```

# Helper functions

### Format columns

Standardize all annotations across multiple annotators and keep only specific columns

```{r}
format_columns <- function(df, keep_columns)
{
  colnames(df) <- tolower(colnames(df))
  df <- df[,keep_columns]
  
  return(df)
}
```

### Calculating metrics

Create a function that outputs accuracy, cohen's kappa, f1 score, precision and recall between pairs of annotators:

1.  Standardized annotation dataframes for each annotator (dlist)
2.  The "theme" label (frame in this case) to compare annotations across
3.  Rater names A1 and A2
4.  Column names where the labels for A1 and A2 lie respectively
5.  Indicator on what to do with NA values (dropNA = T by default)

```{r}
calculate_scores <- function(dlist, A1, A2, column_A1, column_A2, column_match = "stories_id", dropNA = T, silent = T)
{
  dfA1 <- dlist[[A1]]
  dfA2 <- dlist[[A2]]

  kdf1 <- merge(dfA1, dfA2, by.x = column_match, by.y = column_match, all = F)

  #if column names for A1 and A2 are same
  if(column_A1 == column_A2)
  {
    kdf <- kdf1[,c(paste0(column_A1, ".x"), paste0(column_A2, ".y"))]
  }

  else
  {
    kdf <- kdf1[,c(column_A1, column_A2)]
  }

  colnames(kdf) <- c("A1", "A2")

  if(dropNA)
  {
    kdf <- kdf[complete.cases(kdf),]
  }

  klist <- kappa2(kdf) #kappa calculation
  k <- round(klist$value, 2) #Cohen's Kappa

  acc <- round(accuracy(kdf[,1],kdf[,2]),2) #accuracy
  prec <- round(precision(kdf[,1], kdf[,2]), 2) #precision
  rec <- round(recall(kdf[,1], kdf[,2]), 2) #recall
  F1 <- round(2*prec*rec/(prec + rec), 2) #F1 score cos the f1 func is broken

  metric_vec <- c(acc, k, F1, prec, rec)
  
  if(silent == F)
  {
    print(paste0("Accuracy is ", acc, "; Cohen's kappa is ", k, "; F1 score is ",F1,
               "; Precision is ", prec, "; Recall is ", rec))
  }

  return(metric_vec)
}

```

# Relevance

### Compile

Annotations available:

1.  First author (A1) manual annotations for rounds 1 and 2
2.  Third author (A2) manual annotations for round 1 and 2
3.  Naive Bayes
4.  BERT large cased
5.  DeBERTa-v3-small
6.  Llama-3.1-8B-Instruct (8-bit quantized) zero-shot
7.  Llama-3.1-8B-Instruct (8-bit quantized) fine-tuned
8.  Claude Sonnet 4
9.  GPT-OSS-20B

```{r}

#create a list of dataframes
relevance_list <- list(
  #First author
  "a1_r1" = read.csv("Relevance/Human/A1_r1.csv"),
  "a1_r2" = read.csv("Relevance/Human/A1_r2.csv"),
  
  #Second author
  "a2_r1" = read.csv("Relevance/Human/A2_r1.csv"),
  "a2_r2" = read.csv("Relevance/Human/A2_r2.csv"),
  
  #Smaller language models
  "naive_bayes" = read.csv("Relevance/SLM/naive_bayes_test.csv"),
  "bert" = read.csv("Relevance/SLM/bert_test.csv"),
  "deberta" = read.csv("Relevance/SLM/deberta_test.csv"),
  
  #Large language models (post 2023)
  "llama_zero-shot" = read.csv("Relevance/LLM/Predicted/llama3.1:8b-instruct-q8_0_zshot_withCB.csv"),
  "llama_fine-tuned" = read.csv("Relevance/LLM/Predicted/llama3.1:8b-instruct-q8_0_ft.csv"),
  "claude_zero-shot" = read.csv("Relevance/LLM/Predicted/claude-sonnet-4-20250514_zshot_withCB.csv"),
  "gpt-oss_zero-shot" = read.csv("Relevance/LLM/Predicted/gpt-oss-20b-zshot_withCB.csv")
)

keep_columns <- c("stories_id","media_name", "title","url","text","keywords", "relevance")
relevance_list <- lapply(relevance_list, format_columns, keep_columns = keep_columns)

save(relevance_list, file = "Labelled/Relevance_test.RData")

```

### Evaluate

Calculate how different annotators performed on the relevance coding task with respect to each other on the test set.

```{r}
A1 <- "a1_r1" #annotator 1
A2 <- "a2_r1" #annotator 2

column_A1 <- "relevance" #predictions for annotator 1
column_A2 <- "relevance" #predictions for annotator 2
column_match <- "stories_id" #column used to track the annotation file IDs


calculate_scores(relevance_list, A1, A2, column_A1, column_A2, column_match)

```

### Extrapolate

Extrapolate labels onto the entire dataset of 3314 articles for further analysis.

```{r}
df_all <- read.csv("Relevance/SLM/bert_all.csv")
write.csv(df_all, "Labelled/relevance_all.csv", row.names = F)

```

# Frame Detection

## Compile

Annotations available:

1.  First author (A1) manual annotations for rounds 1 and 2
2.  Third author (A2) manual annotations for round 1 and 2
3.  Naive Bayes
4.  BERT large cased
5.  DeBERTa-v3-small
6.  Naive Bayes with class imbalance management
7.  BERT large cased with class imbalance management
8.  DeBERTa v3 small with class imbalance management
9.  Llama-3.1-8B-Instruct (8-bit quantized) zero-shot
10. Llama-3.1-8B-Instruct (8-bit quantized) fine-tuned
11. Llama-3.1-8B-Instruct (8-bit quantized) with LLM as judge
12. Llama-3.1-8B-Instruct (8-bit quantized) with human feedback
13. Llama-3.1-8B-Instruct (8-bit quantized) with decision trees
14. Llama-3.1-8B-Instruct (8-bit quantized) with confidence scores
15. Claude Sonnet 4
16. GPT-OSS-20B

```{r}
#create a list of dataframes
framing_list <- list(
  #First author
  "a1_r1" = read.csv("Framing/Human/A1_r1.csv", check.names = F),
  "a1_r2" = read.csv("Framing/Human/A1_r2.csv", check.names = F),
  
  #Second author
  "a2_r1" = read.csv("Framing/Human/A2_r1.csv", check.names = F),
  "a2_r2" = read.csv("Framing/Human/A2_r2.csv", check.names = F),
  
  #Smaller language models
  "naive_bayes" = read.csv("Framing/SLM/Predicted/Test/naive_bayes_test.csv", check.names = F),
  "bert" = read.csv("Framing/SLM/Predicted/Test/bert_test.csv", check.names = F),
  "deberta" = read.csv("Framing/SLM/Predicted/Test/deberta_test.csv", check.names = F),
  
  
  #class imbalanace managed
  "nb_imb" = read.csv("Framing/SLM/Class_imbalance/NB_test.csv", check.names = F),
  "bert_imb" = read.csv("Framing/SLM/Class_imbalance/BERT_test.csv", check.names = F),
  "deberta_imb" = read.csv("Framing/SLM/Class_imbalance/DeBERTa_test.csv", check.names = F),

  #Large language models (post 2023)
  "llama_zero-shot" = read.csv("Framing/LLM/Predicted/Test/llama3.1:8b-instruct-q8_0_zshot_1to1.csv", check.names = F),
  "llama_ft" = read.csv("Framing/LLM/Predicted/Test/llama3.1:8b-instruct-q8_0_ft_1to1.csv", check.names = F),
  "llama_llm_judge" = read.csv("Framing/LLM/Predicted/Test/llama3.1:8b-instruct-q8_0_LLM_judge.csv", check.names = F),
  "llama_human_feedback" = read.csv("Framing/LLM/Predicted/Test/llama3.1:8b-instruct-q8_0_human_judge.csv", check.names = F),
  "llama_decision" = read.csv("Framing/LLM/Predicted/Test/llama3.1:8b-instruct-q8_0_decision_tree.csv", check.names = F),
  "llama_confidence" = read.csv("Framing/LLM/Predicted/Test/llama3.1:8b-instruct-q8_0_confidence.csv", check.names = F),
  "claude_zero-shot" = read.csv("Framing/LLM/Predicted/Test/claude-sonnet-4-20250514_zshot.csv", check.names = F),
  "gpt-oss_zero-shot" = read.csv("Framing/LLM/Predicted/Test/gpt-oss-20b-zshot.csv", check.names = F)
)

keep_columns <- c("stories_id","media_name", "title","url","text","keywords",
                  "sexual stigma and transmission routes",
                  "racial disparities and stigmatising name",
                  "global relations",
                  "public health failure",
                  "epidemic preparedness and surveillance",
                  "human-interest stories",
                  "broader health issues"
                  )
framing_list <- lapply(framing_list, format_columns, keep_columns = keep_columns)

save(framing_list, file = "Labelled/Framing_test.RData")
```

## Evaluate

Calculate how different annotators performed on the relevance coding task with respect to each other on the test set.

```{r}
A1 <- "deberta" #annotator 1
A2 <- "deberta_imb" #annotator 2

column_match <- "stories_id" #column used to track the annotation file IDs

frames <- c("sexual stigma and transmission routes",
            "racial disparities and stigmatising name",
            "global relations",
            "public health failure",
            "epidemic preparedness and surveillance",
            "human-interest stories",
            "broader health issues")


for(f in frames)
{
  print(f)
  calculate_scores(framing_list, A1, A2, column_A1 = f, column_A2 = f, silent = F)
}

```

## Extrapolate

Compile all available annotations articles:

Annotations available:

1.  First author (A1) manual annotations for 500 articles
2.  Third author (A2) manual annotations for 100 articles
3.  Naive Bayes for 2221 articles
4.  BERT large cased for 2221 articles
5.  DeBERTa-v3-small for 2221 articles
6.  Claude Sonnet 4 for 2224 articles

```{r}
#create a list of dataframes
framing_list <- list(
  #First author
  "A1" = rbind(read.csv("Framing/Human/A1_r2.csv", check.names = F), read.csv("Framing/Human/A1_train.csv", check.names = F)),
  
  #Second author
  "A2" = read.csv("Framing/Human/A2_r2.csv", check.names = F),
  
  #Smaller language models
  "naive_bayes" = read.csv("Framing/SLM/Predicted/All/nb_all.csv", check.names = F),
  "bert" = read.csv("Framing/SLM/Predicted/All/bert_all.csv", check.names = F),
  "deberta" = read.csv("Framing/SLM/Predicted/All/deberta_all.csv", check.names = F),

  #Large language models (post 2023)
  "llama_zero-shot" = read.csv("Framing/LLM/Predicted/All/llama3.1:8b-instruct-q8_0_zshot_all.csv", check.names = F),
  "claude_zero-shot" = read.csv("Framing/LLM/Predicted/All/claude-sonnet-4-20250514_zshot_all.csv", check.names = F)
)

keep_columns <- c("stories_id","media_name", "title","url","text","keywords",
                  "sexual stigma and transmission routes",
                  "racial disparities and stigmatising name",
                  "global relations",
                  "public health failure",
                  "epidemic preparedness and surveillance",
                  "human-interest stories",
                  "broader health issues"
                  )

framing_list <- lapply(framing_list, format_columns, keep_columns = keep_columns)

save(framing_list, file = "Labelled/Framing_all.RData")
```
